---
output: github_document
---



```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# MQmetrics

<!-- badges: start -->
<!-- badges: end -->

The goal of MQmetrics is to analyze Proteomics data from LC-MS/MS. It takes the output tables from MaxQuant and plots multiple parameters.

## Installation


You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("BioAlvaro/MQmetrics")
```
## Example

MQmetrics it's easy to use, and will provide you information about your Proteomics MaxQuant analysis.

Start by loading the library. By default, the MaxQuant results will be stored in a folder named *Combined*. The directory to that folder is all you need to use MQmetrics.

```{r example}
library(MQmetrics)

MQPathCombined <- '/home/alvaro/Documents/MaxQuant/example5/combined/'
```

The main function of the package is **generateReport()**. It will generate a PDF report containing several visualizations and tables from different MaxQuant output tables.
As input it is only necessary to provide the Path to the *Combined* folder of MaxQuant output.

```{r generate_reprort,  eval = FALSE}
generateReport(MQPathCombined, long_names = TRUE, sep_names = '_')
```

Two useful parameters of every function including **generateReport()** are:
*long_names* and *sep_names*. They will allow a clear visualization of those samples that have long names separated by a character. In this example, the Experiment names are one full string separated by underscores (_): *PLK010_QC02_210121_HeLa_125ng_150meth*. 

```{r example_long_names, message=FALSE, warning=FALSE}
# ReadDataFromDir reads all the files needed from the MaxQuant output and remove
# Potential contaminants, reverse, and proteins identified by site only.
files <- ReadDataFromDir(MQPathCombined, remove_contaminants = TRUE) 

PlotIntensity(files[["proteinGroups.txt"]], long_names = TRUE, sep_names = '_')
```

Check the vignettes for more information.


