---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# MQmetrics

<!-- badges: start -->
<!-- badges: end -->

The goal of MQmetrics is to analyze Proteomics data from LC-MS/MS. It takes the output tables from MaxQuant and plots multiple parameters.

## Installation


You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("BioAlvaro/MQmetrics")
```
## Example

MQmetrics will easily generate a report containing all the info with one function.
For that it is needed to provide the directory of the combined folder of MaxQuant output.

```{r example, eval = FALSE}
library(MQmetrics)

MQPathCombined <- '/home/alvaro/Documents/MaxQuant/combined/'

generateReport(MQPathCombined, long_names = TRUE, sep_names = '_')

```

One of the most useful parameters of every function including **generateReport()** are:
*long_names* and *sep_names*. They will allow a clear visualization of sample that have long names separated by a character. Example:



